### 2024-09-26 paper/paper.qmd
Edited first part of data analysis

---
title: "Investigating Bicycle Theft in Toronto: Trends and Patterns"
author: 
  - Taejun Um
thanks: "Code and data are available at: https://github.com/tonytto/Investigating-Bicycle-Theft-in-Toronto"
date: today
date-format: long
abstract: "In urban environments like Toronto, bicycle theft is as common as in any major city worldwide. This paper examines the trends and patterns of bicycle theft in Toronto from 2014 to 2022, focusing on the cost of bicycles and the types of locations where the thefts occurred. Data from this period show that the overall number of bicycle thefts did not significantly decline but began to decrease gradually from 2020, particularly in thefts occurring at residential properties and those involving high-end bicycles. These findings can inform community efforts and assist Toronto's public authorities in developing more effective strategies to protect citizens' property and reduce bicycle theft."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)
library(knitr)

data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

```

# Introduction

Toronto is the most populous and densely populated city in Canada. As a result, traffic congestion makes it difficult for people to navigate the city, and many citizens report this as a significant issue [@traffic]. This has led to an increase in the number of cyclists using bicycles for efficient point-to-point travel within the city. Currently, approximately 659 km of cycling routes are interwoven like a spider's web throughout Toronto [@bikelane]. Unfortunately, bicycle thefts remain a persistent problem for residents.

This paper examines reported bicycle thefts in Toronto from 2014 to 2022. The main objective is to observe and analyze the number of thefts, the locations of the incidents, and the price range of the stolen bicycles. Specifically, we will compare trends in the locations of incidents by year alongside changes in the price range of the targeted bicycles. Through this investigation, I confirmed that thefts steadily increased from 2014 until 2020, when the number of thefts began to decline. Additionally, there have been notable changes in the locations where thefts occurred and the price range of the stolen bicycles over the years. Understanding these trends and patterns will help prevent future bicycle thefts, and inform new measures by Toronto’s public authorities to enhance theft prevention.

The remainder of this paper is structured as follows: @sec-data discusses the data and process of data cleaning, and @sec-first-point discusses our findings, limitation of the study, and suggestions to reduce bicycle theft.

# Data {#sec-data}
## Overview {#sec-data-overview}

The dataset used in the paper covers bicycle thefts reported to the Toronto Police system between 2014 and 2024. This dataset has been updated as of June 2024 and is considered reliable as it is based on the Toronto Police system.

The dataset was downloaded from Toronto's Open Data Portal using opendatatoronto package [@opendatatoronto]. Its setup was completed through R [@citeR] with the help of tidyverse package [@citetidyverse], janitor package [@citejanitor] and knitr package [@citeknitr].

The original dataset contained 28 variables, but I used the following: ID, occurrence date, occurrence year, occurrence month, occurrence hour, premises type, and bike cost. Then, I created a new group called "cost group" based on the bike cost, categorizing the bicycles according to their price range. Since the dataset is based on the date the theft was reported, I organized the data by the actual date of occurrence to reflect events from 2014 to 2022, excluding incidents that occurred before 2014. Additionally, during the data cleaning phase, observations with missing information were excluded. As a result, the dataset used in this paper contains 29,217 observations. (@tbl-head) shows an first 5 row of the data used.

```{r}
#| label: tbl-head
#| tbl-cap: First 5 rows of data
#| warning: false
#| message: false
#| echo: false

data |>
  slice(1:5) |>
  kable(
    digits = 4,
    booktabs = TRUE,
    linesep = ""
  )
```

## Number of Occurrences by Time

First, we will examine how many bicycle thefts occurred by year. In @fig-occur-year-1, the line chart records thefts yearly and the table records the number of incidents in numbers. Bicycle thefts showed an increasing trend from 2014, peaking in 2018 with 3,418 recorded incidents, followed by a noticeable decrease starting in 2021. The decrements that happened in 2021 continued through 2023. According to @fig-occur-year-2, the number of incidents recorded in 2022 was 2,491 which was fewer than 2,511 incidents recorded in 2014. This indicates that the trend of bicycle theft has been decreasing over the last three years.

```{r}
#| label: fig-occur-year
#| fig-cap: Number of bicycle theft incidents by year
#| warning: false
#| message: false
#| echo: false

date_counts <- data |>
  group_by(occ_year) |>
  summarise(total_occurrences = n())

ggplot(date_counts, mapping = aes(x = occ_year, y = total_occurrences)) +
  geom_line() +
  geom_point(size = 2) +
  theme_minimal() +
  labs(x = "Year", y = "Number of Occurrences")

```
```{r}
#| label: tbl-occur-year
#| tbl-cap: Number of bicycle theft incidents by year
#| warning: false
#| message: false
#| echo: false

date_counts |>
  slice(1:10) |>
  kable(
    col.names = c("Year", "Number of Occurrences"),
    digits = 4,
    booktabs = TRUE,
    linesep = ""
  )
```

@fig-month shows the number of incidents that occurred by month with each line representing a different year. This graph indicates there are clear differences by season. Each year, there was a sharp increase in thefts during the middle of the year, while the beginning and end of the year saw a sharp decline, showing a cyclical pattern. The number of incidents peaked in the summer months (June to August) and the incidents in January and December dropped to less than one-third compared to the summer months in all years.

```{r}
#| label: fig-month
#| fig-cap: Number of bicycle theft incidents by month
#| warning: false
#| message: false
#| echo: false
#| fig-pos: "H"

data$occ_month <- 
  factor(data$occ_month, levels = c("January", "February", "March", "April", "May", "June", 
                                        "July", "August", "September", "October", "November", "December"))

month_grouped <- data |>
  group_by(occ_year, occ_month) |>
  summarise(number_of_observations = n(), .groups = "drop")

ggplot(month_grouped, mapping = aes(x = occ_month, y = number_of_observations, color = factor(occ_year), group = occ_year)) +
  geom_line(size = 0.5) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Month", y = "Number of Occurrences", color = "Year") +
  theme_minimal() +
  theme(text = element_text(size = 12), axis.text.x = element_text(angle = 90)) +
  scale_color_brewer(palette = "Set3")

```

### By hour

@fig-hour represents the number of bicycle theft incidents over an hour. @fig-hour-1 shows the average number of incidents across all years, while the @fig-hour-2 illustrates the number of incidents for each year over an hour. Graph 1 shows that there is a steady increase in thefts from 5 AM to 6 PM, and after that, the numbers begin to decline. Notably, there are three distinct peaks at 9 AM, 12 PM, and 6 PM, corresponding to periods of high social activity and movement. On average over the past decade, the lowest number of incidents occurred at 5 AM, with 37.3 thefts reported annually, whereas 6 PM saw the highest, with 218 thefts per year. From 1 AM to 7 AM, fewer than 100 incidents were recorded each year, but beyond that, the number consistently exceeded 100.

```{r}
#| label: fig-hour
#| fig-cap: Number of bicycle theft incidents by hour in 10 years
#| fig-subcap: ["Average in Year","Separated in Each Year"]
#| warning: false
#| message: false
#| echo: false
#| layout-ncol: 2
#| fig-pos: "H"

time_grouped <- data |>
  group_by(occ_hour) |>
  summarise(number_of_observations = n()/10)

ggplot(time_grouped, mapping = aes(x = occ_hour, y = number_of_observations)) +
  geom_line(size = 0.5) +
  geom_point(size = 2) +
  labs(x = "Hour", y = "Number of Occurrences") +
  theme_minimal()

time_grouped_sep <- data |>
  group_by(occ_year, occ_hour) |>
  summarise(number_of_observations_h = n(), .groups = "drop")

ggplot(time_grouped_sep, mapping = aes(x = occ_hour, y = number_of_observations_h, color = factor(occ_year), group = occ_year)) +
  geom_line(size = 0.5) +
  geom_point(size = 2) +
  scale_color_brewer(palette = "Set1") +
  labs(x = "Month", y = "Number of Occurrences", color = "Year") +
  theme_minimal() +
  scale_color_brewer(palette = "Set3")
```

## Types of Crime Scenes

In this dataset, crime scenes have been categorized into seven types: Apartment, Commercial, Educational, House, Outside, Transit, and Others. As seen in @fig-crime-scenes, there has been significant fluctuation in crimes committed in residential areas. Residential areas are classified into two categories: apartments and houses. Among them, apartment-related thefts showed a steady linear increase until 2019, but in 2020, the number dropped by around 200 incidents in one year, returning to 2017 levels by 2022. House-related thefts followed a zigzag pattern, increasing from 2014 and reaching a peak in 2020, before dropping to less than half by 2022.

Another interesting point is that thefts committed outdoors were more frequent than those in apartments before 2019. However, starting in 2019, apartment thefts surpassed outdoor thefts, until 2022, when outdoor thefts once again became more common than apartment thefts.

```{r}
#| label: fig-crime-scenes
#| fig-cap: Number of occurrences categorized by crime scene
#| echo: false
#| warning: false
#| message: false
#| fig-pos: "H"

data |>
  ggplot(mapping = aes(x=occ_year, fill = premises_type)) + 
  geom_bar(position = "dodge2") +
  facet_wrap(~ premises_type) +
  labs(x = "Year", y = "Number of Occurrences", fill = "Types of Crime Scenes") +
  theme(
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.4, "cm"),legend.position = "bottom"
    )

```

## Cost of Stolen Bicycle

The original dataset includes a variable for the price of stolen bicycles, but it does not contain a variable for grouped prices. Therefore, I created a new variable called cost_group by grouping the prices in increments of $500. @fig-stolen-bike shows a bar graph with number of bike stolen by year categorized by cost group. From 2020 to 2022, the thefts of relatively inexpensive bicycles, priced under $1,000, have significantly decreased. On the other hand, high-end bicycles priced over $2,000 were stolen about 250 times in 2014. But this number has steadily increased, reaching approximately 500 thefts by 2022.
```{r}
#| label: fig-stolen-bike
#| fig-cap: Total cost of bicycle by year
#| fig-pos: "H"
#| echo: false
#| warning: false
#| message: false

data <-
  data|>mutate(
  cost_group =
      factor(cost_group, levels = c("< $500", "$500 < $1000", "$1000 < $1500", "$1500 < $2000", "$2000 <")))
data |>
  ggplot(mapping = aes(x=occ_year, fill = cost_group)) + 
  geom_bar() +
  facet_wrap(~ cost_group) +
  labs(x = "Year", y = "Cost of Bicycle", fill = "Cost Group") +
  theme(
    legend.text = element_text(size = 7),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.4, "cm"),
    legend.position = "bottom"
  )
```

# Discussion

## Overall {#sec-first-point}
The analysis of bicycle theft in Toronto has revealed trends and patterns over the nine years of theft incidents. All trends shifted starting in 2020, which appears to be a result of the lockdowns due to the pandemic. First, the overall number of thefts began to decrease from 2020, indicating a significant reduction in outdoor activities. Bicycle owners were riding their bikes less frequently, and thieves were also unable to engage in outdoor activities due to the lockdowns.

One of the most notable trends is that the daily theft graph exhibits a cyclical pattern. This indicates that most crimes occur in the summer, while very few happen in the winter. The winter in Toronto, with its snow and cold weather, makes cycling difficult for both owners and thieves, which likely explains the lack of activity in both groups during this season.

When examining thefts by location, we observed significant changes in residential area thefts. With the increase in remote work and online learning, more people have been at home, leading to a substantial decrease in thefts within residential spaces.

There were also intriguing observations regarding the price range of stolen bicycles. Year after year, the number of thefts of more expensive bicycles has been on the rise. This can be interpreted in two ways: either bicycle thieves have started to specifically target high-end bicycles, or the prices of bicycles available on the market have increased. What is clear is that the rising theft rates of expensive bicycles have led to more severe property losses for citizens. Bicycle theft may be considered a minor issue, but it is a clear crime that deprives someone of their valuable property or an essential means of transportation for their life.

## Implication

According to this data, the Toronto Police should strengthen patrols and response efforts during the summer to prevent bicycle thefts. As of the data up to 2022, thefts in residential areas have significantly decreased; however, with the end of the lockdown, there is a possibility that these incidents may rise again. Therefore, it would be beneficial to conduct preventive campaigns targeting residential areas.

Regardless of their price, all bicycles are important, but there has been a recent trend of increasing thefts of high-end bicycles. To protect citizens' property, it is essential to deliver messages of caution to the community.

## Weaknesses and next steps

This dataset has a significant amount of missing data, and unfortunately, I had to drop some parts of the dataset. Additionally, only theft incidents that were reported are included, meaning that incidents that were not reported for various reasons are not accounted for. These missing pieces of information may have influenced the trends or patterns identified in this paper.

\newpage

\appendix

# Appendix {-}

\newpage

# References